Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-08-12 doc/report/msc-report.tex
data model.

% TODO: renskriv.
The typical use case for Zohmg is to read one or more column-qualifiers,
which correspond to one or more points in n-space, over a range of time.
Therefor, the keys are made up of the time and unit of the measurement,
and the qualifier queried for represents the point in n-space. More on
this in the data model section.

Explain our data model, the relation between column-families, column-qualifiers
and projections and dimensions.

@@

% TODO: data interchange.
The reducer needs to communicate to the OutputReader under what row key,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form \\
\texttt{\{"column-family:qualifier" : value\}}.

@@

% TODO: not truly correct
% TODO: A detailed explanation of generally everything in the darling dir.
The Darling output format for HBase. Between Dumbo, which Zohmg uses for data
import, and Hadoop JSON is used for serializing data.

>>>>

The underlying data store is HBase, which is a column-oriented database. Column
families are pre-defined, but everything else, such as column qualifiers and
rowkeys, can assume arbitrary values.

Zohmg models the aggregated data with this in mind. An HBase table is created
for each data set, and in the table is a single column family called 'unit'. For
every unit that the data set contains there will be a column qualifier. For
example, 'unit:pageviews' would be one such column qualifier in the example
above. Employing the qualifier to encode arbitrary unit names means that the
user can add new units to his data set configuration without having to change
the schema of the HBase table.

Zohmg stores one or more projections of the data, and each projection is made up
of one or more dimensions in addition to the ever-present time dimension. Every
dimension is a tuple of dimension name and dimension value. The dimension tuples
of the projection are stored in the rowkey. The rowkey is a string; the
dimension names and values are separated by a dash. As such, the dash is one of
the few disallowed characters in dimension names and values.

For example, a rowkey for the projection of the user agent and status code
dimensions might look like this:

\begin{lstlisting}[caption=Example of rowkeys,captionpos=b]
    useragent-firefox-statuscode-200-20090420
    useragent-firefox-statuscode-200-20090421
    useragent-firefox-statuscode-200-20090422
    ..
    useragent-firefox-statuscode-404-20090420
    useragent-firefox-statuscode-404-20090421
    useragent-firefox-statuscode-404-20090422
    ..
    useragent-safari-statuscode-200-20090420
    useragent-safari-statuscode-200-20090421
    ..
\end{lstlisting}

The rowkeys are sorted lexicographically. This means that all entries for the
user agent 'firefox' will be found close to each other. If the data analyst is
more interested in finding all entries with status code 200, it would make sense
to set up the projection so that status code is the first dimension.

@@

The reducer needs to communicate to the OutputReader under what row key,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form \\
\texttt{\{"column-family:qualifier" : value\}}.

@@

% TODO: not truly correct
% TODO: A detailed explanation of generally everything in the darling dir.
The Darling output format for HBase. Between Dumbo, which Zohmg uses for data
import, and JSON is used for serializing data.
